---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

Loading all libraries

```{r, echo=FALSE}
library(tidyverse)
library(sjmisc)
library(ggplot2)
library(skimr)
library(kableExtra)
```

```{r, echo=FALSE}
f <- "https://raw.githubusercontent.com/difiore/ada-2022-datasets/main/KamilarAndCooperData.csv"
d <- read_csv(f, col_names = TRUE)
head(d)
s <- skim(d) # formats results to a wide table
# here we make use of the `%>%` operator and {dplyr} verbs... see below
s %>%
  filter(skim_variable == "Scientific_Name" | skim_type == "numeric") %>%
  rename(variable=skim_variable, missing=n_missing, mean=numeric.mean,
    sd=numeric.sd, min=numeric.p0, p25=numeric.p25, median=numeric.p50,
    p75=numeric.p75, max=numeric.p100, hist=numeric.hist) %>%
  select(variable, missing, mean, sd, min, median, max, hist) %>%
  # drop p25 and p75 for purposes of display
  kable() %>%
  kable_styling(font_size = 10)
detach(package:skimr)
detach(package:kableExtra)
```

Challenge 1: Making boxplots of log(female body mass) ~ family

```{r}
boxplot(log(d$Body_mass_female_mean)~d$Family)

ggplot(data = d, aes(x=Family, y=log(Body_mass_female_mean)))+
  geom_boxplot(na.rm=TRUE)
```

Challenge 2: plot the relationship between female body size and female brain size.

```{r}

plot(x=d$Brain_Size_Female_Mean, y=d$Body_mass_female_mean)
plot(x=log(d$Brain_Size_Female_Mean), y=log(d$Body_mass_female_mean))

ggplot(data=d, aes(x=Brain_Size_Female_Mean, y=Body_mass_female_mean))+
  geom_point(na.rm = TRUE)
```

Challenge 3: Build your own bivariate scatterplot of log(MaxLongevity_m) by log(Body_mass_female_mean) using the “KamilarAndCooperData” dataset.

```{r}
ggplot(data=d, aes(x=log(MaxLongevity_m), y=log(Body_mass_female_mean)))+
  geom_point(na.rm=TRUE)+
  geom_smooth(method = "lm", color="blue")

```

Challenge 4: Select the variables Brain_Size_Female_Mean, Body_mass_female_mean, MeanGroupSize, WeaningAge_d, MaxLongevity_m, HomeRange_km2, and DayLength_km from data frame d and plot scatterplots of all pairs of variables.

```{r}
d1= d %>% 
  select(Brain_Size_Female_Mean, Body_mass_female_mean, MeanGroupSize, WeaningAge_d, MaxLongevity_m, HomeRange_km2, DayLength_km)
pairs(data=d1, ~.)
```


In class dplyr challenge

```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2022-datasets/main/KamilarAndCooperData.csv"
d <- read_csv(f, col_names = TRUE)
d2 = d %>% 
  mutate("Binomial" =paste(Genus, Species, sep=",")) %>% 
  select(Binomial, Family, Genus, Brain_Size_Species_Mean, Body_mass_male_mean, Body_mass_female_mean) %>% 
  group_by(Family) %>% 
  summarise(avg1 = mean(Brain_Size_Species_Mean, na.rm = TRUE),
            avg2 = mean(Body_mass_male_mean, na.rm = TRUE),
            na.rm = TRUE) %>% 
  arrange(desc(avg1))
d2
```
